\clearpage
\section{IRAF Statistics}
\index{statistics}

The basic approach to statistics with IRAF is a bit different from what you
will see in textbooks. Most textbooks in statistics refer to farm animals,
crops etc. Here we are dealing with a mix of photon arrival statistics
(Poisson) together with things that 


\subsection{Gaussian Distribution}

Also called the ``normal distribution''. \index{Gaussian distribution}


\begin{subequations}
\begin{align}
f(x) &= \frac{1}{\sqrt{2\pi\sigma^2} } e^{-\frac{(x-\mu)^2}{2\sigma^2}}
\end{align}
\end{subequations}

where:  $\mu$ is the mean, and $\sigma$ is the variance. The distribution is
``normalized'' by subtracting from each observation the true mean, or the expected
value.

\clearpage
In general the terms used are:

\begin{subequations}
\begin{align}
    \textrm{mean} &= \bar{x} \;=\; \frac{1}{N}\sum_{1}^{N} x_{i} \\
\textrm{with:~ ~ ~}  y = x - \bar{x} \notag \\
\textrm{variance} &= \frac{1}{(N-1)}\;\sum_{1}^{N} y_{i}^{2} \\
  \textrm{stddev} &= \sqrt{\textrm{variance}} \\
    \textrm{skew} &= \frac{1}{(N-1)}\;\sum_{1}^{N} \frac{y_{i}}{\sigma}^{3} \\
\textrm{kurtosis} &= \left(  \frac{1}{(N-1)}\;\sum_{1}^{N} \frac{y_{i}}{\sigma}^{4} \right) -3
\end{align}
\end{subequations}

The ``mode'' \cite{wiki-stetson} is the most frequent value in a
dataset. This is hard to determine when high-precision floating point
numbers are used. A little algebra on Pearson's use gives the form
promoted by Stetson and the one that appears in IRAF.

\begin{subequations}
\begin{align}
mode &= 3\times(median) - 2\times (mean),
\end{align}
\end{subequations}

In IRAF and as promoted by Peter Stetson the values are not 'precise'.
One gets  used to these things.

The "median" is the "middle value" of a sorted set. If the count of
items in the set are odd this results in a clear winner. If the count
is even, the dividing point falls between two value -- the
\dhl{WARNING} average is returned.


\index{statistics!mean}\index{statistics!skew}\index{statistics!stddev}
\index{statistics!variance}\index{statistics!kurtosis}

There are cases where the statistics are more important, and a task will
use a more customized approach.

\textbf{\emph{Help:}} imstatistics

\subsection{Mode}

Mode is the most frequent value. What does that mean with discrete
noisy floating point numbers? The common way, in counting statistics,
is to bin find the bin with the most counts. It is often the case with
sensors where the distribution is NOT
Unimodal\index{statistics!unimodal} but the superposition of two
distributions. This can be seen with basic histograms where the shape
is not symmetric.

Stetson, in some places, uses Pearson's trick to computes the mode.
For unimodal distributions, $(2\bar(x) - \rm{mode})/3$. 

When Iraf wants a better estimate, it fits a 2nd order polynomial to
the data, determines the peak and reports that value.


\subsection{SNR Statistics T.B.D.}

T.B.D.

\subsubsection{Photometry Statistics T.B.D.}
T.B.D.

\subsubsection{Spectroscopy Statistics T.B.D.}
T.B.D.



